# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].


This is an other form of reference a markdown table in R book down>

| Feature | Count | Miss | Card |          Mode         |  MF |   MP   |    Mode2   | M2F |  M2P   |
|:-------:|:-----:|:----:|:----:|:---------------------:|:---:|:------:|:----------:|:---:|:------:|
|         | 338   | 0    | 2    | Femenino              | 188 | 55.62% | Masculino  | 150 | 44.38% |
|         | 338   | 1    | 5    | Docente               | 143 | 42.31% | Empleado   | 70  | 20.71% |
|         | 338   | 0    | 4    | Más mental            | 156 | 46.15% | Mental     | 143 | 42.31% |
|         |       |      |      | que físico            |     |        |            |     |        |
|  DD4    | 338   | 2    | 5    | ASD                   | 150 | 44.38% | Católica   | 129 | 38.17% |
|         | 338   | 0    | 5    | Casada(o)             | 181 | 53.55% | Soltera(o) | 119 | 35.21% |
|         | 338   | 0    | 5    | Ninguno               | 284 | 84.02% | Otro       | 26  | 7.69%  |
|         | 338   | 1    | 5    | No                    | 272 | 80.47% | Otra       | 34  | 10.06% |

Table: (\#tab:demographic-features) Demographic features in dataset


The table \@ref(tab:demographic-features) shows the demographic features for the dataset that will be analyzed



